\documentclass{article}

% page layout
\usepackage[letterpaper, textwidth=125mm, textheight=230mm]{geometry}
\usepackage{setspace}
\setstretch{1.08}
\sloppy\sloppypar\raggedbottom\frenchspacing

% list management
\usepackage{enumitem}
\setlist{itemsep=0.0ex}
\setlist[enumerate]{label=\textsl{(\alph*)}}

% problem typesetting -- THIS IS SO INSANE
\usepackage{amsmath, amsthm}
\makeatletter
\def\th@definition{%
  \thm@notefont{}% same as heading font
  \normalfont % body font
}
\makeatother
\usepackage{etoolbox}
\makeatletter
\patchcmd{\@thm}{\thm@headpunct{.}}{\thm@headpunct{:}}{}{}
\makeatother
\theoremstyle{definition}
\newcommand{\problemname}{Problem}
\newtheorem{problem}{\problemname}
\newcommand{\probref}[1]{\problemname~\ref{#1}}
\theoremstyle{remark}
\newcommand{\notename}{Note}
\newtheorem{note}{\notename}[problem]
\newcommand{\noteref}[1]{\textit{\notename}~\ref{#1}}

% math macros
\newcommand{\dd}{\mathrm{d}}

\begin{document}
\section*{Crowded-field and confusion problem set}
\noindent
\textit{by} \textbf{David W. Hogg}

\medskip\noindent
There are some disagreements between Hogg and Casey about what is and isn't possible in modeling crowded fields.
I (Hogg) am talking out of my ass; this problem set is designed to make my conjectures explicit, and to test them.

\begin{problem}[power-law flux distributions]
In terms of \emph{apparent flux} $S$, sources tend to be drawn from power-laws of the form
\begin{align}
    \frac{\dd\ln n}{\dd\ln S} = A - \alpha\,\ln\frac{S}{S_0} ~,
\end{align}
where $A$ is a log amplitude, $\alpha$ is a power law exponent, and $S_0$ is a characteristic flux.
Write code to draw flux values $S$ from the implied probability distribution, and demonstrate by making plots that your code is correct.
Your code should take as input the power-law exponent $\alpha$, a flux interval $[S_{\min}, S_{\max}]$ flux in which to work, to and the number $n$ of stars you want returned in that interval.

Now what exponent $\alpha$ do you expect for bright stars in the nearby Galaxy?
Make the assumption that stars uniformly fill three-dimensional space and that the luminosity distance and volume elements are trivially Euclidean.

What exponent $\alpha$ do you expect for the stars in a globular cluster, at the faint end, roughly?

How do your answers transform if you ask about the log number per \emph{magnitude} instead of log flux?
\end{problem}

\begin{problem}[naive detectability of an isolated star]\label{prob:naive}
Imagine that you have a background-limited image taken with an ideal instrument consisting of a square grid of identical pixels.
Background-limited, here, means that every pixel has a noise contribution that doesn't depend significantly on the sources that contribute to that pixel's intensity.
Imagine that the pixel-convolved point-spread function (PRF) in this image is precisely Gaussian and has a full width at half maximum (FWHM) $w=3.0$\,pixels.
Let's imagine that we took an image that contains, within its field of view, exactly one unresolved star of flux $S$, and the star is nowhere near the edge of the imaging field of view.
Finally, assume that there is a white (uncorrelated, iid) noise contribution to every pixel with value $\sigma=1.$ (in whatever units).
Assume that there is perfect and uniform detector sensitivity.
At what flux $S_5$ is this star measured at $5$-sigma?
Equivalently, what is the expected uncertainty $\sigma_S$ on the brightness of this star?

Give your answers in terms of the noise level $\sigma$ and the PRF FWHM $w$.
That is, I gave values for $\sigma$ and $w$ but give your answers symbolically.
Get as precise or as approximate as you like; discuss the subtle issues.

Extra credit: What is the expected uncertainty on the \emph{position} (within the image, in pixels) of the star?
Give your answer in terms of $S$, $\sigma$, and $w$.
\end{problem}

\begin{note}
The pixel-convolved point-spread function, or, in astronomer parlance, PRF, is the thing that can be purely \emph{sampled} (that is, not \emph{integrated}) at the pixel-center locations.
\end{note}

\begin{problem}[Generate crowded fields]
Make a $100\times 100$ pixel grid and fill it with stars.
Here are the requirements:
Generate stellar fluxes $S$ with the power law $\alpha=1$ (the barely diverging, barely Olbers-paradox case).
Generate stellar positions uniformly in the image, but generate in a larger region, so the stars extend naturally beyond the edges of the image.
Generate stars in the flux interval $[S_{\min},S_{\max}]=[0.003\,S_5, 3.0\,S_5]$ where $S_5$ is the flux level you computed in \probref{prob:naive}.
That is, generate stars down to extremely low fluxes such that the faintest would be ``detected'' at the 0.015-sigma level (of course they won't be detected).
Generate enough stars that there are of order 300 stars at true fluxes $S>S_5$ within the image.
Make a synthetic image using the PRF stated in \probref{prob:naive} and add uniform white noise as required by \probref{prob:naive}.
Keep (don't delete) the true catalog that went into the image!
Deliver this image and this catalog, and code that can generate arbitrary numbers of new examples.
\end{problem}

\begin{note}
Why $300$? Because that corresponds to about 30 beams per source.
\end{note}

\begin{note}
You will have to generate a \emph{fuck ton} of stars to meet the requirements of this problem.
\end{note}

\begin{problem}
Use whatever standard astrophysical code you would use to detect stars in the generated image.
Use the best (frequentist, standard) crowded-field code to which you have access; don't skimp here.
Ask the code to deliver all stars down to the flux level $S_5$ you computed in \probref{prob:naive}.
Document your choices.
Compare this catalog to the expectations you generated in \probref{prob:naive}.
Comment.
\end{problem}

\end{document}
